---
title: 'Distribution and bioecology of Auchenorrhyncha (Hemiptera) in western Andalusian agroecosystems, with data for the control of potential vectors of Xylella fastidiosa'  
subtitle: "Chapter 4: Auchenorrhyncha in the Agroecosistems of Western Andalusia with focus on the vectors of *Xylella fastidiosa*: community description, seasonal abundance, and effects of site-specific, climatic and landscape factors"
author: "Laura Avivar-Lozano, José Mª. Molina-Rodríguez , Sergio Pérez-Guerrero"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
    number_section: TRUE
    theme: "cosmo"
    code_download: true
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
  body {
    text-align: justify;
  }
</style>
```

```{r message=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  skimr, writexl, readr, readxl, data.table,
  tidyverse, tidylog, summarytools, here,
  plotrix, lubridate, textshape, ggrepel,ggforce,
  bipartite, FSA, vegan, mvabund, reshape2, iNEXT,indicspecies
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(
  dev        = "png",
  fig.path   = paste0(here::here("figure", "C4_Multivariante"), "/plot-"),
  fig.width  = 10,
  fig.height = 8, 
  dpi=300
)
```

# Auchenorrhyncha Woody Crops

**Do Auchenorrhyncha communities vary across different woody crops?**

-   Prepare data

```{r}
tax<-read_excel(here("data/resultado_taxonomia_final2.xlsx")) 

tax1<-tax%>% dplyr::filter(año!=2020, !iden %in% c("noiden", "genero"))

```

-   Create an interaction matrix for each infraorder

```{r}
tax2<-tax1%>% dplyr::filter(suborden=="Cicadomorpha")%>% #Redo with 'Fulgoromorpha'
  group_by(crop, subfamily) %>% 
  summarise(ind=sum(total, na.rm = TRUE)) %>%  
  ungroup() %>%  
  pivot_wider( names_from=subfamily,
               names_sep=".", 
               values_from=ind, 
               values_fn=sum, values_fill=0) %>%  
  ungroup()

tax2<-textshape::column_to_rownames(tax2, loc =1)
tax2 <- as.matrix(tax2)# transform the data into a matrix

ourweb<-tax2

#Vectors specifying the colors for the high trophic level, interactions and lower trophic level

paleta_suave <- colors_cultivos <- c( "lightpink",   "dodgerblue",   "orange",   "olivedrab",    "purple")

high.colors <- c(rep("deeppink4",94))

low.colors <- rep(paleta_suave, times = c(1, 1, 1,1,1))  

int.colors <- rep (c (rep("dimgray",500)))
```

## Interaction network analysis

```{r}
plotweb (ourweb, method="normal", #other is method="cca"
         arrow="center",
         labsize= 0.5,
         text.rot="90", 
         col.low= low.colors,
         col.high= high.colors,         
         col.interaction= int.colors,
         bor.col.low="transparent",  
         bor.col.high="transparent",    
         bor.col.interaction="transparent" )
```

## Alpha diversity indices

### Rarefaction curves

```{r}
species<-dcast(data = tax1, formula = crop ~ code,
               fun.aggregate = sum, value.var = "total")
species<-species[,1:96]
num_muestreos <- data.frame(
  crop = c("Almond", "Olive", "Vineyard", "Citrus", "Blueberry"),
  Num_sample = c(204, 794, 302, 210, 284))
species <- left_join(species, num_muestreos, by = "crop")
abundancia_matriz <- as.matrix(species[, -c(1, ncol(species))])
rownames(abundancia_matriz) <- species$crop
mincrop<-204
abundancia_normalizada <- abundancia_matriz / species$Num_sample
abundancia_estandarizada <- abundancia_normalizada * mincrop
abundancia_estandarizada <- round(abundancia_estandarizada)
abundancia_estandarizada <- t(abundancia_estandarizada)
```

```{r}
rare_data <- iNEXT(as.data.frame(abundancia_estandarizada), 
                   q = 0, datatype = "abundance")
```

-   Plot

```{r}
ggiNEXT(rare_data, type = 1) + 
  labs(title = "Rarefaction Curves",
       x = "Standardized sample size",
       y = "Richness") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1", name = "Crop") +  
  scale_fill_brewer(palette = "Set1", guide = "none") +  
  guides(shape = "none") +  
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.major = element_line(size = 0.1),
        panel.grid.minor = element_blank()) +
  geom_line(linewidth = 0.2)
```

-   Community matrix and metadata

```{r}
species<-dcast(data = tax1, formula = fecha+  crop ~ code,
               fun.aggregate = sum, value.var = "total")
sum(rowSums(species[, -c(1, 2,98)]) == 0)
species <- species[rowSums(species[, -c(1, 2,98)]) > 0, ] #delete 58 lines
sp<-species[,3:97] #Community matrix 
species$crop<-as.factor(species$crop)
sp_c<-species[,2] #metadata
sp_c <- as.data.frame(sp_c)
colnames(sp_c) <- "crop" 
sp_c$crop<-as.factor(sp_c$crop)
```

### Diversity indices

-   Calculation of diversity indices

```{r}
diversity_alfa<-sp |> 
  as.data.frame() |> 
  mutate(Richness = specnumber(sp),
    Shannon = diversity(sp, index = "shannon"),
    Simpson = diversity(sp, index = "simpson"),
    Pielou = Shannon / log(Richness)) |>  #  Pielou's Evenness 
  bind_cols(sp_c)
```

-   Plot diversity

```{r}
diversity_alfa_long <- diversity_alfa %>%
  pivot_longer(cols = c(Richness, Shannon, Simpson, Pielou),
    names_to = "Indice",
    values_to = "Valor")
diversity_alfa_long$Indice <- factor(diversity_alfa_long$Indice, 
                                      levels = c("Richness", "Shannon", "Simpson", "Pielou"))
diversity_alfa_long <- diversity_alfa_long %>% filter(!is.na(Valor))
```

-   Plot

```{r}
ggplot(diversity_alfa_long, aes(x = crop, y = Valor, fill = crop)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1, outlier.alpha = 0.8) +
  geom_jitter(width = 0.1, alpha = 0.5, size = 0.5) +
  facet_wrap(~ Indice, scales = "free_y", ncol = 2) +
  theme_minimal(base_size = 16) +
  labs(title = "",
       x = "",
       y = expression (alpha~"Diversity indices")) +
  # Eliminar la leyenda
  theme(legend.position = "none",
        # Ajustar el fondo y las líneas de la cuadrícula para un aspecto más limpio
        panel.grid.major = element_line(color = "grey90"),
        panel.grid.minor = element_blank(),
        # Alinear mejor las etiquetas del facet
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(size = 14, 
                                   color = "black",
                                   face = "bold",
                                   angle = 45, 
                                   hjust = 1))
```

-   Non-parametric analyses Kruskal-Wallis test

    ```{r}
    K.S<-kruskal.test(Richness ~ crop, data = diversity_alfa)
    print(K.S)
    K.H<-kruskal.test(Shannon ~ crop, data = diversity_alfa)
    print(K.H)
    K.D<-kruskal.test( Simpson ~ crop, data = diversity_alfa)
    print(K.D)
    K.J<-kruskal.test( Pielou ~ crop, data = diversity_alfa)
    print(K.J)
    ```

-   Dunn's test for the results of pairwise comparisons between crops
    for the alpha diversity indices, only if Kruskal-Wallis is
    significant.

    ```{r}
    dunn_S <- dunnTest(Richness ~ crop, data = diversity_alfa, method = "bonferroni")
    print(dunn_S)

    dunn_H <- dunnTest(Shannon ~ crop, data = diversity_alfa, method = "bonferroni")
    print(dunn_H)

    dunn_D <- dunnTest(Simpson ~ crop, data = diversity_alfa, method = "bonferroni")
    print(dunn_D)
    ```

## NDMS and PERMANOVA

```{r}
tax<-read_excel(here("data/resultado_taxonomia_final2.xlsx")) 
tax1<-tax%>% dplyr::filter(año!=2020, !iden %in% c("noiden", "genero"))
```

Variation in community composition by crop and sampling date using NMDS

-   Species abundance matrix by crop and sampling date for NMDS analysis

```{r}
species <-dcast(data = tax1, formula = fecha+ crop ~species, fun.aggregate = sum, value.var = "total") 
sum(rowSums(species[-c(1, 2,98)]) == 0) 
species <- species[rowSums(species[, -c(1,2,98)])> 0, ]
```

```{r}
sp<-species[,3:97] #Community matriz (Abundance)
sp_crop <- as.factor(species$crop) #metadata crop
```

### NDMS

```{r,results = "hide"}
sp1<-metaMDS(sp,autotransform = FALSE,
             distance = "bray", 
             k = 2, 
             trymax = 999)
```

```{r}
print(sp1)
```

-   Fitting species to NMDS and extracting species vectors

```{r}
# Ajustar especies al NMDS
species_fit <- envfit(sp1, sp, permutations = 999)
species_scores <- as.data.frame(scores(species_fit, "vectors"))
species_scores$species <- rownames(species_scores)  # Asignar nombres de especies
```

```{r}
species_scores$r2 <- species_fit$vectors$r
species_scores <- species_scores[species_fit$vectors$pvals < 0.05 & species_scores$r2 > 0.2, ] # Filtrar solo especies con alta significancia
```

-   Plot

```{r}
nmds_points <- as.data.frame(scores(sp1, display = "sites"))
nmds_points$crop <- sp_crop  # add crop
```

```{r}
ggplot(data = nmds_points, aes(x = NMDS1, y = NMDS2, color = crop)) +
  geom_point(size = 3, alpha = 0.8) +  # Puntos de los sitios
  geom_mark_ellipse(aes(fill = crop), alpha = 0.2) +  # Elipses por cultivo
  geom_segment(data = species_scores, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.2, "cm")), color = "black", inherit.aes = FALSE) +  
  geom_text_repel(data = species_scores, aes(x = NMDS1, y = NMDS2, label = species),
                  color = "black", size = 3, inherit.aes = FALSE) +  
  scale_shape_manual(values = c(16, 17, 15)) +  # Formas: círculo, triángulo, cuadrado
  theme_minimal()
```

### PERMANOVA

```{r}
bray_dist <- vegdist(sp, method = "bray")
permanova<- adonis2(bray_dist ~ sp_crop,
                    permutations = 999)
print(permanova)
```

#### Dispersion

Verification of the assumption of homogeneous dispersion in PERMANOVA

```{r}
b_c <- betadisper(vegdist(sp, method = "bray"), group = sp_crop)
permutest(b_c, permutations = 999)
boxplot(b_c)
```

A priori, the Auchenorrhyncha community is similar across different
crops. However, since some differences have been found in the diversity
indices, we proceed to identify the indicator species for each crop and
habitat. The models will be used to further deepen the analysis.

## IndVal_index_Auchenorryncha

```{r}
matriz_abundancia <- tax1 %>%
  group_by(fecha, id_finca, crop, species) %>%
  summarise(ind = sum(total, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = species, values_from = ind, values_fill = 0) %>%
  select(-fecha, -id_finca)
matriz_abundancia <- as.data.frame(matriz_abundancia)
matriz_abundancia[, -1] <- apply(matriz_abundancia[, -1], 2, as.numeric)
```

```{r}
indval_result <- multipatt(matriz_abundancia[, -1], matriz_abundancia$crop, func = "IndVal.g", control = how(nperm = 999))
summary(indval_result)
```

-   FDR Correction for Multiple Testing Analysis

```{r}
indval_result$sign$p_value_fdr <- p.adjust(indval_result$sign$p.value, method = "fdr")
signif_species <- indval_result$sign[indval_result$sign$p_value_fdr < 0.05, ] %>%
  rownames_to_column(var = "Species") %>%
  na.omit()
```

```{r}
long_data <- signif_species %>%
  pivot_longer(cols = starts_with("s."), names_to = "crop", values_to = "Presencia") %>%
  filter(Presencia == 1) %>%
  group_by(Species) %>%
  mutate(max_stat = max(stat)) %>%
  ungroup() %>%
  mutate(Species = reorder(Species, -max_stat),
         crop = gsub("^s\\.", "", crop),
         significance = case_when(
           p_value_fdr <= 0.001 ~ "***",
           p_value_fdr <= 0.01 ~ "**",
           p_value_fdr <= 0.05 ~ "*",
           TRUE ~ ""
         ))
```

-   Heatmap "IndVal Auchenorryncha Crops"

```{r}
# Crear el heatmap (MATRIZ DE COLOR) con los nombres de las especies en cursiva y los asteriscos
ggplot(long_data, aes(x = crop, y = Species, fill = stat)) +
  geom_tile(color = "white") +  
  geom_text(aes(label = significance), color = "black", size = 5, vjust = 0.5) +  # Añadir asteriscos de significancia
  scale_fill_gradient(low = "lightblue", high = "blue", na.value = "grey90") +  
  labs(title = "IndVal Auchenorryncha Crops",
       x = "",
       y = "",
       fill = "IndVal index") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1, size = 12, color="black"), 
        axis.text.y = element_text(size = 12, face = "italic",color="black"),  
        axis.title = element_text(size = 12),  
        plot.title = element_text(size = 14, face = "bold"), 
        panel.grid = element_blank(),  
        panel.spacing = unit(0.5, "lines"))  
```

-   Adapt the code to extract the Heatmap 'IndVal Auchenorryncha
    Habitat"

# Auchenorrhyncha Olive

**How do Auchenorrhyncha communities vary across different olive grove
plots?**

-   Data

```{r}
tax3<-tax1 |> dplyr::filter(crop=='Olive')
```

## Alpha diversity indices.

### Rarefaction curves

-   Matrix

```{r}
species<-dcast(data = tax3, formula = fecha+id_finca~ code,
                fun.aggregate = sum, value.var = "total")
sp<-species[,3:72] #Community matriz (Abundance)
sp_plot <- as.character(species$id_finca) #metadata
```

```{r}
esfuerzo_minimo <- 138
```

```{r}
Plots <- unique(sp_plot)
```

```{r}
spec_list <- list()
```

```{r}
for (Plot in Plots) {
  subset_abundancia <- sp[which(sp_plot == Plot), , drop = FALSE]
  
  # Calcular specaccum con un máximo de individuos de 'esfuerzo_minimo'
  spec <- specaccum(subset_abundancia, method = "random", permutations = 100)
  
  # Crear un data frame hasta 'esfuerzo_minimo'
  max_sample <- min(length(spec$richness), esfuerzo_minimo)
  
  spec_df <- data.frame(
    Individuos = 1:max_sample,
    Riqueza = spec$richness[1:max_sample],
    SD = spec$sd[1:max_sample],
    Plot = Plot
  )
  
  # Almacenar en la lista
  spec_list[[Plot]] <- spec_df
}
```

```{r}
spec_combined <- bind_rows(spec_list, .id = "Plot")
```

```{r}

# Graficar las curvas de rarefacción con barras de error
ggplot(spec_combined, aes(x = Individuos, y = Riqueza, color = Plot, fill = Plot)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = Riqueza - SD, ymax = Riqueza + SD), alpha = 0.2, color = NA) +
  theme_minimal() +
  labs(title = "Rarefaction curves",
       x = "Standardized sample size",
       y = "Richness") +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

```

### Diversity indices

-   Matrix

```{r}
species<-dcast(data = tax3, formula = fecha+id_finca~ code,
                fun.aggregate = sum, value.var = "total")
sp<-species[,3:72] #Community matriz (Abundance)
sp_plot <- as.factor(species$id_finca) #metadata
sp_plot <- as.data.frame(sp_plot)
colnames(sp_plot) <- "plot"
```

-   Calculation of diversity indices

```{r}
diversity_alfa<-sp |> 
  as.data.frame() |> 
  mutate(Richness = specnumber(sp),
    Shannon = diversity(sp, index = "shannon"),
    Simpson = diversity(sp, index = "simpson"),
    Pielou = Shannon / log(Richness)) |>  #  Pielou's Evenness 
  bind_cols(sp_plot )
```

-   Plot diversity

```{r}
diversity_alfa_long <- diversity_alfa %>%
  pivot_longer(cols = c(Richness, Shannon, Simpson, Pielou),
    names_to = "Indice",
    values_to = "Valor")
diversity_alfa_long$Indice <- factor(diversity_alfa_long$Indice, 
                                      levels = c("Richness", "Shannon", "Simpson", "Pielou"))
diversity_alfa_long <- diversity_alfa_long %>% filter(!is.na(Valor))
```

-   Plot

```{r}
ggplot(diversity_alfa_long, aes(x = plot, y = Valor, fill = plot)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1, outlier.alpha = 0.8) +
  geom_jitter(width = 0.1, alpha = 0.5, size = 0.5) +
  facet_wrap(~ Indice, scales = "free_y", ncol = 2) +
  theme_minimal(base_size = 16) +
  labs(title = "",
       x = "",
       y = expression (alpha~"Diversity indices")) +
  # Eliminar la leyenda
  theme(legend.position = "none",
        # Ajustar el fondo y las líneas de la cuadrícula para un aspecto más limpio
        panel.grid.major = element_line(color = "grey90"),
        panel.grid.minor = element_blank(),
        # Alinear mejor las etiquetas del facet
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(size = 10, 
                                   color = "black",
                                   face = "bold",
                                   angle = 45, 
                                   hjust = 1))
```

-   Non-parametric analyses Kruskal-Wallis test

    ```{r}
    K.S<-kruskal.test(Richness ~ plot, data = diversity_alfa)
    print(K.S)
    K.H<-kruskal.test(Shannon ~ plot, data = diversity_alfa)
    print(K.H)
    K.D<-kruskal.test( Simpson ~ plot, data = diversity_alfa)
    print(K.D)
    K.J<-kruskal.test( Pielou ~ plot, data = diversity_alfa)
    print(K.J)
    ```

-   Dunn's test for the results of pairwise comparisons between crops
    for the alpha diversity indices, only if Kruskal-Wallis is
    significant.

    ```{r}
    dunn_S <- dunnTest(Richness ~ plot, data = diversity_alfa, method = "bonferroni")
    print(dunn_S)

    dunn_H <- dunnTest(Shannon ~ plot, data = diversity_alfa, method = "bonferroni")
    print(dunn_H)

    dunn_D <- dunnTest(Simpson ~ plot, data = diversity_alfa, method = "bonferroni")
    print(dunn_D)

    dunn_J <- dunnTest(Pielou ~ plot, data = diversity_alfa, method = "bonferroni")
    print(dunn_J)
    ```

## NDMS and PERMANOVA

```{r}
species <- tax3 %>%
  group_by(fecha, id_finca, code) %>%
  summarise(abundancia_total = sum(total, na.rm = TRUE), .groups = 'drop')

sp <- species %>%
  pivot_wider(names_from = code, 
              values_from = abundancia_total, 
              values_fill = list(abundancia_total = 0))

sp<-sp %>%dplyr:: select(-3)

sp <- sp[rowSums(sp[, -c(1, 2)]) > 0, ] 
sp_plot <- sp %>%
  select(fecha, id_finca)
sp <- sp %>%
  select(-fecha, -id_finca) %>%
  as.matrix()
```

### NMDS

```{r,results = "hide"}
sp1<-metaMDS(sp,
             autotransform = FALSE, 
             distance = "bray", 
             k = 2, 
             trymax = 999)
```

```{r}
print(sp1)
```

```{r}
coord_muestras <- as.data.frame(sp1$points)
coord_especies <- as.data.frame(wascores(sp1$points,sp))
```

```{r}
coord_muestras <- cbind(coord_muestras, id_finca = sp_plot$id_finca)
coord_muestras$id_finca <- as.factor(coord_muestras$id_finca)
```

```{r, warning= FALSE}
ggplot(coord_muestras, aes(x = MDS1, y = MDS2, color = id_finca)) +
  stat_ellipse(aes(fill = id_finca), alpha = 0.2, geom = "polygon") +
  geom_text_repel(data = coord_especies, aes(x = MDS1, y = MDS2, label = rownames(coord_especies)), 
                  color = "black", size = 3, max.overlaps = 35) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Línea horizontal discontinua
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Línea vertical discontinua
  xlim(-1.5, 0.5) +
  ylim(-1, 1.3) +
  theme_minimal(base_size = 15) +
  theme(panel.background = element_rect(fill = "white", color = NA),  # Fondo blanco
        panel.grid.major = element_line(color = "gray90"),             # Líneas de cuadrícula suaves
        panel.grid.minor = element_blank()) +                         # Sin cuadrícula menor
  labs(title = "NMDS de Cicadomorpha por Plot", x = "NMDS1", y = "NMDS2")
```

### PERMANOVA

```{r}
sp <- data.matrix(sp)
bray_dist <- vegdist(sp, method = "bray")
permanova<- adonis2(bray_dist ~id_finca,data=sp_plot,
                    permutations = 999)
print(permanova)
```

#### Dispersion

Verification of the assumption of homogeneous dispersion in PERMANOVA

```{r}
b_p <- betadisper(vegdist(sp, method = "bray"), group = sp_plot$id_finca)
permutest(b_p, permutations = 999)
boxplot(b_p)
```

```{r}
plot(b_p, hull = TRUE, ellipse = TRUE, label = TRUE, main = "Homogeneity of Dispersions by Type of Plot")
```

Post-hoc tests for PERMANOVA.

```{r}
#Pruebas post-hoc de permanova de calculo manual
fincas <- unique(sp_plot$id_finca)
pares <- combn(fincas, 2, simplify = FALSE)

resultados <- data.frame(
  Finca1 = character(),
  Finca2 = character(),
  F.Model = numeric(),
  R2 = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE
)

for(par in pares){
  finca1 <- par[1]
  finca2 <- par[2]
  
  indices <- which(sp_plot$id_finca %in% c(finca1, finca2))
  
  matriz_subset <- sp[indices, ]
  metadatos_subset <- sp_plot[indices, ]
  
  metadatos_subset$id_finca <- as.factor(metadatos_subset$id_finca)
  
  bray_subset <- vegdist(matriz_subset, method = "bray")
  
  perma <- adonis2(bray_subset ~ id_finca, data = metadatos_subset)
  
  resultados <- rbind(resultados, data.frame(
    Finca1 = finca1,
    Finca2 = finca2,
    F.Model = perma$F[1],
    R2 = perma$R2[1],
    p.value = perma$`Pr(>F)`[1]
  ))
}

# Ajuste de los valores p por el método de Bonferroni
resultados$p.adjusted <- p.adjust(resultados$p.value, method = "bonferroni")

# Visualización de los resultados
print(resultados)

```

A priori, the Auchenorrhyncha community is different in olive plots. The
models will be used to further deepen the analysis.

## IndVal_index_Auchenorryncha

```{r}
matriz_abundancia <- tax3 %>%
  group_by(fecha, id_finca,  species) %>%
  summarise(ind = sum(total, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = species, values_from = ind, values_fill = 0) %>%
  select(-fecha)
matriz_abundancia <- as.data.frame(matriz_abundancia)
matriz_abundancia[, -1] <- apply(matriz_abundancia[, -1], 2, as.numeric)
```

```{r}
indval_result <- multipatt(matriz_abundancia[, -1], matriz_abundancia$id_finca, func = "IndVal.g", control = how(nperm = 999))
summary(indval_result)
```

-   FDR Correction for Multiple Testing Analysis

```{r}
indval_result$sign$p_value_fdr <- p.adjust(indval_result$sign$p.value, method = "fdr")
signif_species <- indval_result$sign[indval_result$sign$p_value_fdr < 0.05, ] %>%
  rownames_to_column(var = "Species") %>%
  na.omit()
```

```{r}
long_data <- signif_species %>%
  pivot_longer(cols = starts_with("s."), names_to = "plot", values_to = "Presencia") %>%
  filter(Presencia == 1) %>%
  group_by(Species) %>%
  mutate(max_stat = max(stat)) %>%
  ungroup() %>%
  mutate(Species = reorder(Species, -max_stat),
         plot = gsub("^s\\.", "", plot),
         significance = case_when(
           p_value_fdr <= 0.001 ~ "***",
           p_value_fdr <= 0.01 ~ "**",
           p_value_fdr <= 0.05 ~ "*",
           TRUE ~ ""
         ))
```

-   Heatmap "IndVal Auchenorryncha Crops"

```{r}
ggplot(long_data, aes(x = plot, y = Species, fill = stat)) +
  geom_tile(color = "white") +  
  geom_text(aes(label = significance), color = "black", size = 5, vjust = 0.5) +  # Añadir asteriscos de significancia
  scale_fill_gradient(low = "lightblue", high = "blue", na.value = "grey90") +  
  labs(title = "IndVal Auchenorryncha Olive",
       x = "",
       y = "",
       fill = "IndVal index") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1, size = 12, color="black"), 
        axis.text.y = element_text(size = 12, face = "italic",color="black"),  
        axis.title = element_text(size = 12),  
        plot.title = element_text(size = 14, face = "bold"), 
        panel.grid = element_blank(),  
        panel.spacing = unit(0.5, "lines"))  
```
