---
title: 'Chapter 4: Auchenorrhyncha in the Agroecosistems of Western Andalusia with focus on the vectors of *Xylella fastidiosa*: community description, seasonal abundance, and effects of site-specific, climatic and landscape factors'
subtitle: "GLMMs"
author: "Laura Avivar-Lozano, José Mª. Molina-Rodríguez , Sergio Pérez-Guerrero"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
    number_section: TRUE
    theme: "cosmo"
    code_download: true
    code_folding: "show"
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
  body {
    text-align: justify;
  }
</style>
```

```{r message=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  skimr, writexl, readr, readxl, data.table,
  tidyverse, tidylog, summarytools, here,
  plotrix, lubridate, textshape, ggrepel,ggforce,patchwork,
  FSA, vegan, mvabund, reshape2, glmmTMB, MuMIn, lmtest, DHARMa, psych, performance, car, emmeans, sjPlot,effects, visreg, parameters, MASS, effectsize, splines, ggplot2, ggeffects, pROC,tweedie, lubridate,bbmle, moments, flextable,TMB, psych
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(
  dev        = "png",
  fig.path   = paste0(here::here("figure", "C4_Modelos"), "/plot-"),
  fig.width  = 10,
  fig.height = 8, 
  dpi=300
)
```

```{r}
au<-read_excel(here("data/octubre_modelos.xlsx")) #matriz adultos con envieroment
data<-read_excel(here("data/matriz_ninfas_modelos.xlsx")) #matriz ninfas envieroment
```

```{r,echo = FALSE,results = "hide"}
##Data estructura adultos
au$id_loc<-as.factor(au$id_loc)
au$id_finca<-as.factor(au$id_finca)
au$crop<-as.factor(au$crop)
au <- au %>%
  mutate(crop = dplyr::recode(as.character(crop),
      "almendro"  = "Almond",
      "arándano"  = "Blueberry",
      "cítricos"  = "Citrus",
      "olivar"    = "Olive",
      "vid"       = "Vineyard"))
au$habitat<-as.factor(au$habitat)

au <- au %>%
  mutate(habitat = dplyr::recode(as.character(habitat),
          "borde"="Edge",
          "copa"="Canopy", 
          "interior"= "Within-crop"))  

au$manejo<-as.factor(au$manejo)
au$seasonality<-as.factor(au$seasonality)
au$year<-as.factor(au$year)
au$logabun<-log(au$abundancia+1)
au$logCV<-log(au$cv+1) 
au$lognp_w<-log(au$np_w+1) 
au$lognp_wo<-log(au$np_wo+1) 
au$lognp_h<-log(au$np_h+1) 
au$lognp_s<-log(au$np_s+1) 
au$lognp_i<-log(au$np_i+1) 
au$lognp_art<-log(au$np_art+1) 
au$logDSH<-log(au$dsh+1)
au$logPp<-log(au$prep_smn+1)
au$logDS<-log(au$dsf+1)
au$logshdi<-log(au$shdi+1)
au$loged_s<-log(au$ed_s+1) 
au$logca_h<-log(au$ca_h+1)
au$logca_s<-log(au$ca_s+1)
au$logTmx<-log(au$Tmax+1)
au %>%  group_by(id_finca) %>% #numero de aspiraciones por finca
  summarise(count_habitat = sum(n_asp, na.rm = TRUE))
n_muestreo1<-  c("ca01" = 204, "ca02" =204, "ca03" = 136, "co01" = 210, "co02" = 210, "co03" = 70, 
                 "hu01" = 148, "hu02" = 136, "hu03" =96, "hu04" = 138, "se01" = 150, "se02" =192)
au$n_muestreo1 <- n_muestreo1[au$id_finca]
au$logn_m1<-log(au$n_muestreo1+1)
au$logn_c<-log(au$n_cultivo+1)
##Data estructura ninfas
data$id_loc<-as.factor(data$id_loc)
data$id_finca<-as.factor(data$id_finca)
data$crop<-as.factor(data$crop)
data$habitat<-as.factor(data$habitat)
data$manejo<-as.factor(data$manejo)
data$seasonality<-as.factor(data$seasonality)
data$year<-as.factor(data$year)
data$logTm<-log(data$Tmax+1)
data$logca_h<-log(data$ca_h+1)
```


```{r}
set.seed(123)
```

```{r}
options(contrasts=c(factor="contr.sum", ordered="contr.poly"))
control.tmb <- glmmTMBControl(optCtrl=list(iter.max=1000, eval.max=1000), parallel=10) 
```

# Models Auchenorrhyncha woody crops 

## Abundance

```{r}
m1<-glmmTMB(formula = abundancia ~ crop + habitat + ns(dia, 4) +
            (1 | id_loc:year) +logCV+ lognp_w + lognp_wo, data = au, family = genpois(link = "log"), 
            ziformula = ~0, dispformula = ~1, offset = logn_m1,
            control = control.tmb, REML = T)
```

## Richness

```{r}
m2<-glmmTMB(formula = nspp ~ crop + habitat + ns(dia, 4) + (1 | id_loc:year)+
            logCV+ lognp_w  + lognp_wo, data = au, family = genpois(link = "log"),
            ziformula = ~0, dispformula = ~1, offset = logn_m1, 
            control = control.tmb, REML = T)
```

```{r}
p1<-plot_model(m1, 
               show.data=FALSE, 
               type="est", 
               ci.lvl=0.95, 
               terms=NULL, 
               show.values=TRUE, 
               value.size = 6, 
               value.offset=0.4, 
               colors =c("brown1","cyan4") ,
               vline.color = "blueviolet")+
  theme_minimal(base_size = 16)+ 
  theme( panel.border = element_rect(color = "black",
                                     fill = NA, 
                                     linewidth = 0.5),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank())

p2<-plot_model(m2, 
               show.data=FALSE, 
               type="est", 
               ci.lvl=0.95, 
               terms=NULL, 
               show.values=TRUE, 
               value.size = 6, 
               value.offset=0.4, 
               colors =c("brown1","cyan4") ,
               vline.color = "blueviolet")+
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
(p1+p2)+ plot_annotation( tag_levels = "A")
```

```{r}
pred1 <- ggpredict(m1, terms = c( "logCV","habitat","crop"))
a1<-plot(pred1)+
  labs(title = "Abundance",
       x = "",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r,fig.width=4, fig.height=4}
pred2 <-ggpredict(m1, terms = c( "dia","habitat"))
plot(pred2)+
  labs(title = "",
       x = "",
       y = "") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m2, terms = c( "logCV","habitat","crop"))
a2<-plot(pred1)+
  labs(title = "Richness",
       x = "LogCV",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r, fig.width=20, fig.height=26}
(a1/a2)+ plot_layout(guides = "collect") &  
  theme(legend.position = "bottom")
```

```{r,fig.width=4, fig.height=4}
pred2 <-ggpredict(m2, terms = c( "dia","habitat"))
plot(pred2)+
  labs(title = "",
       x = "Day",
       y = "") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        legend.position = "bottom")
```

# Models Auchenorrhyncha olive

```{r, echo = FALSE,results = "hide"}
au_oli<-subset(au, crop== "Olive")
au_oli$id_loc<- droplevels(au_oli$id_loc) # Actualiza los niveles del factor id_loc
au_oli$id_finca<- droplevels(au_oli$id_finca)
au_oli %>% dplyr::count(id_finca,n_asp) %>% 
  group_by(id_finca) %>% summarise(nolivar=sum(n))

n_muestreo3<-  c("ca02" =102,  "co01" = 105, 
                 "hu04" = 69, "se01" = 75, "se02" =96)
au_oli$n_muestreo3 <- n_muestreo3[au_oli$id_finca]
au_oli$logn_m3<-log(au_oli$n_muestreo3+1)
```

## Abundance

```{r}
m3<-glmmTMB(formula = abundancia ~ habitat + logCV + 
               ns(dia, 4) + lognp_wo + (1 | id_finca:year), data = au_oli, 
             family = nbinom2(link = "log"), ziformula = ~0, dispformula = ~1, 
             offset = logn_m3, control = control.tmb, REML = T)
```

## Richness

```{r}
m4<-glmmTMB(formula = nspp ~ habitat + logCV +  ns(dia,4) +
              lognp_wo +   (1 | id_finca:year), 
            data = au_oli, 
            family = genpois(link = "log"), 
            ziformula = ~0, 
            dispformula = ~1, 
            offset = logn_m3, 
            control = control.tmb, 
            REML = T)
```

```{r}
p3<-plot_model(m3, 
               show.data=FALSE, 
               type="est",
               ci.lvl=0.95, 
               terms=NULL, 
               show.values=TRUE,
               value.size = 6, 
               value.offset=0.4, 
               colors =c("brown1","cyan4") ,
               vline.color = "blueviolet",
               axis.lim=c(0.01,50))+
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank())

p4<-plot_model(m4, 
               show.data=FALSE, 
               type="est",
               ci.lvl=0.95, 
               terms=NULL, 
               show.values=TRUE,
               value.size = 6, 
               value.offset=0.4, 
               colors =c("brown1","cyan4") ,
               vline.color = "blueviolet",
               axis.lim=c(0.01,50))+
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank())
(p3+p4)+ plot_annotation( tag_levels = "A")
```

```{r}
pred1 <- ggpredict(m3, terms = c( "logCV","habitat"))
g3<-plot(pred1)+
  labs(title = "",
       x = "",
       y = "Abundance") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        legend.position = "bottom")
```

```{r,fig.width=4, fig.height=4}
pred2 <-ggpredict(m3, terms = c( "dia","habitat"))
g4<-plot(pred2)+
  labs(title = "",
       x = "",
       y = "") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m4, terms = c( "logCV","habitat"))
g5<-plot(pred1)+
  labs(title = "",
       x = "LogCV",
       y = "Richness") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        legend.position = "bottom")
```

```{r,fig.width=4, fig.height=4}
pred2 <-ggpredict(m4, terms = c( "dia","habitat"))
g6<-plot(pred2)+
  labs(title = "",
       x = "Day",
       y = "") +
  theme_minimal(base_size = 16)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        legend.position = "bottom")
```

```{r}
(g3+g4)/(g5+g6)+  plot_layout(guides = "collect") &  # Junta las leyendas
  theme(legend.position = "bottom")
       
```
# Adults Aphrophoeidae
```{r,echo = FALSE,results = "hide"}
au_sub <- subset(au, habitat != "Canopy" & id_finca != "co03")
au_sub$habitat<-as.factor(au_sub$habitat)
au_sub$crop<-as.factor(au_sub$crop)

au_sub$id_finca <- droplevels(au_sub$id_finca)
au_sub$crop <- droplevels(au_sub$crop)

au_sub %>% dplyr::count(id_finca,n_asp) #Cambiar el offset
n_muestreo2<-  c("ca01" = 68, "ca02" =68, "ca03" = 68, "co01" = 70, "co02" = 70,  
                 "hu01" = 74, "hu02" = 68, "hu03" =48, "hu04" = 46, "se01" = 50, "se02" =64)
au_sub$n_muestreo2 <- n_muestreo2[au_sub$id_finca]
au_sub$logn_m2<-log(au_sub$n_muestreo2+1)
sum(is.na(au_sub))
```

## Abundance Adults Aphrophoridae (total)
```{r}
m5<-glmmTMB(formula = aphro_total ~  habitat + logCV +ns(dia, 6) +lognp_s + (1 |  id_loc:year),
            data = au_sub, 
            family = genpois(link = "log"), 
            ziformula = ~0, 
            dispformula = ~1, 
            offset = logn_m2, 
            control = control.tmb, 
            REML = T)
```
## Abundance Adults Neophilaenus campestris

```{r}
m6<-glmmTMB(formula = nc ~  logCV + ns(dia, 6) + lognp_s + logDSH + (1 |  id_loc:year),
            data = au_sub,
            family = nbinom2(link = "log"), 
            ziformula = ~0,
            dispformula = ~1,
            offset = logn_m2, 
            control = control.tmb, 
            REML = T)
```

## Abundance Adults Lepyronia coleoptrata
```{r}
m7<-glmmTMB(formula = lc ~ habitat +cv+ logPp + Tmax + lognp_s+(1 | id_loc:year),
            data = au_sub, 
            family = genpois(link = "log"), 
            ziformula = ~0, 
            dispformula = ~1, 
            offset = logn_m2,
            control = control.tmb,
            REML = T)
```

## Abundance Adults Philaenus spumarius

```{r}
#Elimino las localidades que no hay-- cambio data (504 obs)
au_sub1<- subset(au_sub, !(id_loc %in% c("xf03", "xf04", "xf05")))
au_sub1$id_loc<- droplevels(au_sub1$id_loc)
au_sub1$crop<- droplevels(au_sub1$crop)
```

```{r}
m8<-glmmTMB( formula = ps~ cv, 
             data = au_sub1,
             family =nbinom1(link = "log"), 
             ziformula = ~0, 
             dispformula = ~1,
             offset = logn_m2, 
             control = control.tmb, 
             REML =T)
```

```{r}
pred1 <- ggpredict(m5, terms = c( "logCV","habitat"))
b1<-plot(pred1)+
  labs(title = "Adults Aphrophoridae",
       x = "LogCV",
       y = "Abundance") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```
```{r}
pred2 <- ggpredict(m5, terms = c( "dia","habitat"))
b2<-plot(pred2)+
  labs(title = "",
       x = "Day",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r}
(b1+b2)+  plot_layout(guides = "collect") &  # Junta las leyendas
  theme(legend.position = "bottom")
```

```{r}
data<-read_excel(here("data/matriz_ninfas_modelos.xlsx")) #matriz ninfas envieroment
```

```{r,echo = FALSE,results = "hide"}
##Data estructura adultos
au$id_loc<-as.factor(au$id_loc)
au$id_finca<-as.factor(au$id_finca)
au$crop<-as.factor(au$crop)
au <- au %>%
  mutate(crop = dplyr::recode(as.character(crop),
      "almendro"  = "Almond",
      "arándano"  = "Blueberry",
      "cítricos"  = "Citrus",
      "olivar"    = "Olive",
      "vid"       = "Vineyard"))
au$habitat<-as.factor(au$habitat)

au <- au %>%
  mutate(habitat = dplyr::recode(as.character(habitat),
          "borde"="Edge",
          "copa"="Canopy", 
          "interior"= "Within-crop"))  


##Data estructura ninfas
data$id_loc<-as.factor(data$id_loc)
data$id_finca<-as.factor(data$id_finca)
data$crop<-as.factor(data$crop)
data$habitat<-as.factor(data$habitat)
data <- data %>%
  mutate(crop = dplyr::recode(as.character(crop),
      "almendro"  = "Almond",
      "arándano"  = "Blueberry",
      "cítricos"  = "Citrus",
      "olivar"    = "Olive",
      "vid"       = "Vineyard"))
data$habitat<-as.factor(data$habitat)

data <- data%>%
  mutate(habitat = dplyr::recode(as.character(habitat),
          "borde"="Edge",
          "copa"="Canopy", 
          "interior"= "Within-crop"))  
data$manejo<-as.factor(data$manejo)
data$seasonality<-as.factor(data$seasonality)
data$year<-as.factor(data$year)
data$logTm<-log(data$Tmax+1)
data$logca_h<-log(data$ca_h+1)
```


```{r}
set.seed(123)
```

```{r}
options(contrasts=c(factor="contr.sum", ordered="contr.poly"))
control.tmb <- glmmTMBControl(optCtrl=list(iter.max=1000, eval.max=1000), parallel=10) 
```

# Nymphs Aphrophoeidae 
##Abundance total

```{r}
m9<-glmmTMB(formula = ind ~ crop + habitat + manejo + logcv + logTm + logDSH + ns(dia, 4) + lognp_s +
            (1 | id_loc:year), 
            data = data, 
            family = genpois(link = "log"), 
            ziformula = ~0, 
            dispformula = ~1, 
            offset = off_finca, 
            control = control.tmb, 
            REML = T)
```

## Abundance Neophilaenus campestris
```{r}
m10<-glmmTMB(formula = nc~ habitat  + logcv + logPp + logTm + logDSH + ns(dia, 4)+logca_h+ loged_s+lognp_s +lognp_wo + (1 | id_loc:year), 
             data = data,
             family = genpois(link = "log"), 
             ziformula = ~0, 
             dispformula = ~1, 
            offset = off_finca, 
            control = control.tmb, 
            REML = T)
```

## Abundance Lepyronia coleoptrata
```{r}
m11<-glmmTMB(formula = lc ~ habitat +logcv  + ns(dia, 4) + lognp_s + (1 | id_loc:year),
             data = data, 
             family = genpois(link = "log"), 
             ziformula = ~0, 
             dispformula = ~1, 
             offset = off_finca,
             control = control.tmb, 
             REML = T)
```

## Abundance Philaenus spimarius
```{r}
m12<-glmmTMB(formula = ps ~ logcv  + lognp_h + logDSH + (1 | id_loc:year),
             data = data, 
             family = genpois(link = "log"),
             ziformula = ~0, 
             dispformula = ~1,
             offset = off_finca,
             control = control.tmb, 
             REML = T)
```

```{r}
pred1 <- ggpredict(m9, terms = c( "logcv","habitat","crop"))
b3<-plot(pred1)+
  labs(title = "Nymphs Aphrophoridae",
       x = "LogCV",
       y = "Abundance") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r,fig.width=10, fig.height=12}
(b1+b2)/b3+  plot_layout(guides = "collect") &  # Junta las leyendas
  theme(legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m6, terms = c( "dia","logDSH"))
n1<-plot(pred1)+
  labs(title = "",
       x = "Day",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m10, terms = c( "logca_h","logDSH"))
n2<-plot(pred1)+
  labs(title = "",
       x = "Log(ca-H)",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m7, terms = c( "cv","habitat"))
n3<-plot(pred1)+
  labs(title = "",
       x = "CV",
       y = "Abundance") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "bottom")
```

```{r}
pred1 <- ggpredict(m11, terms = c( "lognp_s ","habitat"))
n4<-plot(pred1)+
  labs(title = "",
       x = "Log(np-S)",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "right")
```

```{r, fig.width=4, fig.height=4}
pred1 <- ggpredict(m12, terms = c( "lognp_h ","logDSH"))
n5<-plot(pred1)+
  labs(title = "",
       x = "Log(np-H)",
       y = "") +
  theme_minimal(base_size = 20)+ 
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.2),
        axis.title.x = element_text(),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.text = element_text( color = "black"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
        strip.text = element_text(size = 24, face = "bold"),
        legend.position = "right")
```

```{r, fig.width=10, fig.height=12}

d1<-(n1+n2)+plot_layout(guides = "collect") &  # Junta las leyendas
  theme(legend.position = "right")
d2<-(n3+n4) +plot_layout(guides = "collect") &  # Junta las leyendas
  theme(legend.position = "right")
d1/d2/n5
```

